National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs

### 2020-11-30 Supplement_And_Analysis.Rmd
fixed table small edits

---
title             : "Supplement: Analysis for National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs"
author: "contact: joseph.bulbulia@gmail.com on behalf of Highland et al."
date: "27/11/2020"
authornote: |
  <!-- https://tinyurl.com/ybremelq -->
# keywords          : "keywords"
wordcount         : no
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
# documentclass     : "apa6"
classoption       : "man"
output            : "html_document" 
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# library("papaja")
#r_refs("r-references.bib")
```

@@

# function to clean data, specify number of waves as 3 
d_3<-data_clean_spirit_wellbeing(d,3) # d = dataframe, 3 = number of waves
# function to count uinque ides
show_unique_id(d_3$Id) 
# function to count waves for participants
count_waves_participants(d_3)
# table of beliefs 
table(d_3$Beliefs)
# years Note that we use a continuous response measure particpants. 
range(d_3$years)
```

@@

```{r, include=FALSE}
# must remove Furniture package before running table 1
demographic_table(d_3)
```

```{r, include=FALSE}
beliefs_table(d_3)
```

@@

```{r}
# obtain dag
tidy_ggdag <- dagitty::dagify(
  WB ~ belief + years + partner + nzdep + urban + male + pols + empl,
  WB ~~ partner + years,
  belief ~ years + age + male + ethn,
  partner ~ nzdep + age + belief, 
  nzdep ~ empl,
  pols ~ age  + empl + ethn + years,
  empl ~ nzdep + edu + ethn,
  exposure =  "belief",
  outcome =   "WB")%>%
  tidy_dagitty()

@@

This method reveals two adjustments sets: \{age, employment, male, political conservativism, and time\}, and \{age, ethnicty, male, and time.\}  We report the second set because employment is likely to contain more measurement error: some are not employed because they cannot find employment, others because they are not seeking employment (e.g. retirement).  However we report both model in our frequentist analysis. Neither model choice effects the results. We also report the simple unadjusted model and the fully saturated model (with potential collider and masking effects).

A word of caution, the fully-saturated model -- which would be typical in a longitudional regression analysis -- attenuates the growth in personal well-being and in life satisfaction among those who are skeptical in a spirit or Life-fore.  The inference of this study is not sentitive to this difference because the slopes differ for spirit beliefs across all models. However, we do not wish to infer that skepticism about a spirit or life force entirely prevents growth in personal well-being or life-satisfaction. We advise that researchers do not introduce variables into regression mmodels without systematically. evaluating potential confounding.

@@

Bayesian inference require substantial computational resources. The cost of time is repaid with a model that efficiently pools information to produce a more accurate, and transparent, estimate of uncertainty. The posterior distributions of the model describe intuitive conditional probabilities.

@@

@@

```{r pwi_bayesmodel}
# PWI
mod.1_b <-brm(forumula_bayesPWI, data= d_3, file ="mod.1_b.rds", cores = 8) # note these are very large files
```

```{r lsbayesmodel}
# LIFE SAT
mod.2_b <-brm (forumula_bayesLIFESAT, data= d_3,cores = 8, file ="mod.2_b.rds")
# note these are very large files
```

```{r simple_table, cache =TRUE}
bayestestR::describe_posterior(mod.1_b, ci = 0.90,  test = c("p_direction"),diagnostic = "ESS")
```

```{r simple_table_ls, cache =TRUE}
bayestestR::describe_posterior(mod.2_b, ci = 0.90, test = c("p_direction"), diagnostic = "ESS")
```

@@

```{r Bayesian_model_table, cache=TRUE}
#tab_model(mod.1_b, show.ci= .89) # we really should use something other than 95 but the team consensus is to use 95
tab_model(mod.1_b, mod.2_b, show.ci= .95) # no difference to inference
```

@@

```{r cache =TRUE}
g1dag1<-plot(parameters::model_parameters(mod.1dag))+ ggtitle("Personal well-being Coefficients")
g2dag1<-plot(parameters::model_parameters(mod.2dag))+ ggtitle("Life Satisfaction Coefficients")
```

```{r plot_coefficientplots_LMER_FULL, cache = TRUE}
g1dag1/g2dag1
```

Get the predictions

```{r Bayesian_pwi_predictedSAT, cache=TRUE, dependson= mod.1_b}
get_predictions(mod.1_bDAG)
```

```{r Bayesian_lifesat_predictedSAT, cache=TRUE, dependson= mod.2_b}
get_predictions(mod.2_bDAG)
```

```{r Bayesian_pwi_predictedSAT, cache=TRUE, dependson= mod.1_b}
graph_predictions(mod.1_bDAG, "Predicted Values of Personal Well-Being: Bayesian Model (Confounder Adjusted)")
```

```{r Bayesian_lifesat_predictedSAT, cache=TRUE, dependson= mod.2_b}
graph_predictions(mod.2_bDAG, "Predicted Values of Life Satisfaction: Bayesian Model (Confounder Adjusted)")
```

# describe the posteriors 
```{r simple_table, cache =TRUE}
bayestestR::describe_posterior(mod.1_bDAG, test = c("p_direction"),diagnostic = "ESS")
```

## Test of the Bayesian saturacted model  for the record

```{r setmodelformulaSAT}
# note that BRMS automatically centers so we do not need to center years
forumula_bayesPWI_SATUR <- bf(PWI ~  years * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1+ years|Id))

@@

```{r pwi_bayesmodelSAT}
# PWI
mod.1_bSAT <-brm(forumula_bayesPWI_SATUR, data= d_3, file ="mod.1_bS.rds", cores = 8) # note these are very large files
summary(mod.1_bSAT)
```

```{r Bayesian_pwi_predictedSAT, cache=TRUE, dependson= mod.1_b}
graph_predictions(mod.1_bSAT, "Predicted Values of Personal Well-Being")
```

```{r Bayesian_lifesat_predictedSAT, cache=TRUE, dependson= mod.2_b}
graph_predictions(mod.2_b, "Predicted Values of Life Satisfaction")
```

@@

@@

@@

## Adjusting for Demographics.

@@

## Adjustment using DAG

Here is the table.
```{r tab_dag1, cache = TRUE, dependson=d_3}
sjPlot::tab_model(mod.1dag1,mod.2dag1)
```

```{r Bayesian_lifesat_predict, cache =TRUE}
graph_predictions(mod.1dag1, "Predicted Values of Life Satisfaction: Dag 1")
```
```{r Bayesian_lifesat_predict, cache =TRUE}
graph_predictions(mod.2dag1, "Predicted Values of Life Satisfaction: Dag 1")
```

```{r lmer_model_with_full_covariates, cache =TRUE}
library(lme4)
pwi_formulaDag2 <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C  + Employed + Male + Deprivation_S + Political_Conservativism_S + (1 + yearsC|Id)'
ls_formulaDag2 <- 'LIFESAT ~ yearsC * Beliefs + Employed +  Age_in_Decades_C  + Employed + Male + Deprivation_S + Political_Conservativism_S + (1 + yearsC|Id)'

```{r tab_dag2, cache = TRUE, dependson=d_3}
sjPlot::tab_model(mod.1dag2,mod.1dag2)
```

```{r dag2_plotcreate cache =TRUE}
g1dag2<-plot(parameters::model_parameters(mod.1dag2))+ ggtitle("Personal well-being Coefficients")
g2dag2<-plot(parameters::model_parameters(mod.2dag2))+ ggtitle("Life Satisfaction Coefficients")
```

```{r Bayesian_lifesat_predict, cache =TRUE}
graph_predictions(mod.1dag2, "Predicted Values of Life Satisfaction: Dag 2")
```

```{r Bayesian_lifesat_predict, cache =TRUE}
graph_predictions(mod.2dag2, "Predicted Values of Life Satisfaction: Dag 2")
```

```{r lmer model centering years}
library(lme4)
pwi_formulaC <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1+ yearsC|Id)'

Here is the code for the coefficient plots of our preferred simple model.

@@

Let's plot all the models. We find that uncertainty increases, but that, again, the inference remains robust.

@@

We reran the imputed data analysis.

```{r cache = TRUE, imputedsimplePWI, cache=TRUE}
# run models iterating over imputed data
mod_imputed_pwi<-loop_lmer_model(pwi_formula_simple, imputed3)
```

@@

```{r cache = TRUE}
# run models iterating over imputed data
mod_imputed_ls<-loop_lmer_model(ls_formula_simple,imputed3)

>>>>

---
title             : "Supplement: Analysis for National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs"
author: "contact: joseph.bulbulia@gmail.com on behalf of Highland et al."
date: "27/11/2020"
authornote: |
  <!-- https://tinyurl.com/ybremelq -->
# keywords          : "keywords"
wordcount         : no
bibliography      : ["b-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
# documentclass     : "apa6"
classoption       : "man"
output            : "html_document"

editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library("papaja")
r_refs("b-references.bib")
```

@@

# function to clean data, specify number of waves as 3 
d_3<-data_clean_spirit_wellbeing(d,3) # d = dataframe, 3 = number of waves
# function to count uinque ides
show_unique_id(d_3$Id) 
# function to count waves for participants
count_waves_participants(d_3)
# table of beliefs 
table(d_3$Beliefs)
# years Note that we use a continuous response measure particpants. 
range(d_3$years)
# html talbe
demographic_table(d_3)
```

Sample details are as follows:
```{r sample details}
sample_report(d_3)
```

<!-- ```{r  badlatextable} badd tables -->
<!-- library("arsenal") -->
<!-- summary(arsenal::tableby(Wave ~ Age + -->
<!--                    NZdep + -->
<!--                    Education + -->
<!--                    Employed + -->
<!--                    Ethnic_Categories + -->
<!--                    Male  + -->
<!--                    Has_Partner  + -->
<!--                    Political_Conservativism + -->
<!--                    Urban + -->
<!--                    LIFESAT + PWI + includeNA(Beliefs), data = d_3,digits = 2, total=FALSE, format.p = FALSE, digits.p = 0))%>% -->
<!--     kable( booktabs = T, "latex") -->
<!-- ``` -->

```{r latextable, cache=TRUE}
library(modelsummary)
ts<-d_3 %>%
  dplyr::select(Age, 
                NZdep,
                Education,
                Employed,
                Ethnic_Categories,
                Male,
                Has_Partner, 
                Political_Conservativism,
                Urban, 
                Beliefs, 
                LIFESAT,
                PWI,
                Wave)
modelsummary::datasummary_balance( ~ Wave, data=ts, dinm=FALSE, output = 'table.tex')
# https://github.com/IndrajeetPatil/ggstatsplot
# https://www.datanovia.com/en/lessons/compute-summary-statistics-in-r/
```

@@

```{r demographic table, cache=TRUE}
# must remove Furniture package before running table 1
demographic_table(d_3)
#demographic_table_latex() not wurking  d
```

<!-- ```{r, include=FALSE} -->
<!-- beliefs_table(d_3) -->
<!-- ``` -->

@@

```{r}
# obtain dag
tidy_ggdag <- dagify(
  WB ~ belief + years + partner + nzdep + urban + male + pols + empl,
  WB ~~ partner + years,
  belief ~ years + age + male + ethn,
  partner ~ nzdep + age + belief, 
  nzdep ~ empl,
  pols ~ age  + empl + ethn + years,
  empl ~ nzdep + edu + ethn,
  exposure =  "belief",
  outcome =   "WB")%>%
  tidy_dagitty()

@@

This method reveals two adjustments sets: \{age, employment, male, political conservativism, and time\}, and \{age, ethnicty, male, and time.\}  We report the second set because employment is likely to contain more measurement error: some are not employed because they cannot find employment, others because they are not seeking employment (e.g. retirement).  However we report both model in our frequentist analysis. Neither model choice affects the results. We also report the simple unadjusted model and the fully saturated model (with potential collider and masking effects).

Importantly, in the fully-saturated model -- which would be typical in a longitudinal regression analysis -- attenuates the growth in personal well-being and in life satisfaction among those who are skeptical in a spirit or Life-fore.  The inference of this study is not sensitive to this difference because the slopes differ for spirit beliefs across all models. However, we do not wish to infer that skepticism about a spirit or life force entirely prevents growth in personal well-being or life-satisfaction. We advise that researchers do not introduce variables into regression models without systematically. evaluating potential confounding.

@@

@@

Here we describe the posteriors for both models 
```{r simple_table, cache =TRUE}
bayestestR::describe_posterior(mod.1_bDAG, test = c("p_direction"),diagnostic = "ESS")
```

<!-- ```{r} -->
<!-- cite_r("b-references.bib") -->
<!-- ``` -->

Next we plot the graphs. First we extract the predictions

```{r Bayesian_pwi_predictedSAT_DAG_GET, cache=TRUE}
get_predictions(mod.1_bDAG)
```

```{r Bayesian_lifesat_predictedSAT_DAG_GET, cache=TRUE}
get_predictions(mod.2_bDAG)
```
```{r DAGS, cache =TRUE}
g1dag1<-plot(parameters::model_parameters(mod.1_bDAG))+ ggtitle("Personal well-being Coefficients")
g2dag2<-plot(parameters::model_parameters(mod.2_bDAG))+ ggtitle("Life Satisfaction Coefficients")
```

```{r plot_coefficientplots_LMER_FULL_DAG, cache = TRUE}
g1dag1/g2dag1
```

Here are the coefficient plots

```{r Bayesian_pwi_predictedSAT_DAG, cache=TRUE}
graph_predictions(mod.1_bDAG, "Predicted Values of Personal Well-Being: Bayesian Model (Confounder Adjusted)")
```

```{r Bayesian_lifesat_predictedSAT_DAG, cache=TRUE}
graph_predictions(mod.2_bDAG, "Predicted Values of Life Satisfaction: Bayesian Model (Confounder Adjusted)")
```

## Bayesian model, no confounders addressed

In the simple model, the in

@@

```{r pwi_bayesmodel_DAG}
# PWI
mod.1_b <-brm(forumula_bayesPWI, data= d_3, file ="mod.1_b.rds", cores = 8) # note these are very large files
```

```{r lsbayesmodel_DAG}
# LIFE SAT
mod.2_b <-brm (forumula_bayesLIFESAT, data= d_3,cores = 8, file ="mod.2_b.rds")
# note these are very large files
```

```{r simple_table_DAG, cache =TRUE}
bayestestR::describe_posterior(mod.1_b, ci = 0.90,  test = c("p_direction"),diagnostic = "ESS")
```

```{r simple_table_ls_DAG, cache =TRUE}
bayestestR::describe_posterior(mod.2_b, ci = 0.90, test = c("p_direction"), diagnostic = "ESS")
```

@@

```{r Bayesian_model_table_SIMPLE, cache=TRUE}
#tab_model(mod.1_b, show.ci= .89) # we really should use something other than 95 but the team consensus is to use 95
tab_model(mod.1_b, mod.2_b, show.ci= .95) # no difference to inference
```

@@

## Test of the Bayesian saturated model

```{r setmodelformulaSAT_FULL}
# note that BRMS automatically centers so we do not need to center years
forumula_bayesPWI_SATUR <- bf(PWI ~  years * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1+ years|Id))

@@

```{r pwi_bayesmodelSAT_FULL}
# PWI
mod.1_bSAT <-brm(forumula_bayesPWI_SATUR, data= d_3, file ="mod.1_bS.rds", cores = 8) # note these are very large files
summary(mod.1_bSAT)
```

@@

```{r Bayesian_pwi_predictedSAT, cache=TRUE}
graph_predictions(mod.1_bSAT, "Predicted Values of Personal Well-Being")
```
```{r Bayesian_lifesat_predictedSAT, cache=TRUE}
graph_predictions(mod.2_b, "Predicted Values of Life Satisfaction")
```

@@

## Adjusting for confounds

Here is the table.
```{r tab_dag1, cache = TRUE}
sjPlot::tab_model(mod.1dag1,mod.2dag1)
```

```{r grapwi_predict_DAG, cache =TRUE}
graph_predictions(mod.1dag1, "Predicted Values of Life Satisfaction: Dag 1")
```
```{r gralifesat_predict_DAG, cache =TRUE}
graph_predictions(mod.2dag1, "Predicted Values of Life Satisfaction: Dag 1")
```

```{r lmer_model_with_full_covariates_DAG, cache =TRUE}
library(lme4)
pwi_formulaDag2 <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C  + Employed + Male + Deprivation_S + Political_Conservativism_S + (1 + yearsC|Id)'
ls_formulaDag2 <- 'LIFESAT ~ yearsC * Beliefs + Employed +  Age_in_Decades_C  + Employed + Male + Deprivation_S + Political_Conservativism_S + (1 + yearsC|Id)'

```{r tab_dag2, cache = TRUE}
sjPlot::tab_model(mod.1dag2,mod.1dag2)
```

```{r dag2_plotcreate, cache =TRUE}
g1dag2<-plot(parameters::model_parameters(mod.1dag2))+ ggtitle("Personal well-being Coefficients")
g2dag2<-plot(parameters::model_parameters(mod.2dag2))+ ggtitle("Life Satisfaction Coefficients")
```

```{r pwi_predict_DAG, cache =TRUE}
graph_predictions(mod.1dag2, "Predicted Values of Life Satisfaction: Dag 2")
```

```{r glifesat_predict_DAG, cache =TRUE}
graph_predictions(mod.2dag2, "Predicted Values of Life Satisfaction: Dag 2")
```

```{r lmer model centering years_SATURATED}
library(lme4)
pwi_formulaC <- 'PWI ~  yearsC * Beliefs +  Age_in_Decades_C + Deprivation_S + Education_S  + Employed + Ethnic_Categories + Male  + Has_Partner  + Political_Conservativism_S + Urban + (1+ yearsC|Id)'

@@

## Fully saturated model

@@

@@

The inference is robut even to the poorly specified models using REML estimation.

@@

We reran the imputed data analysis. (Note we ajust for confounds)

```{r cache = TRUE, imputedsimplePWI, cache=TRUE}
# run models iterating over imputed data
mod_imputed_pwi<-loop_lmer_model(pwi_formulaDag1, imputed3)
```

@@

```{r cache = TRUE}
# run models iterating over imputed data
mod_imputed_ls<-loop_lmer_model(ls_formulaDag1 ,imputed3)
